Regionally Additive Models: Explainable-by-design models minimizing feature interactions

### 2023-06-13 report/report.tex
continued algorithm

\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{natbib}
\usepackage{algorithmic}
\bibliographystyle{plainnat}
% set path for bibliography

@@

\begin{algorithm}
\caption{Regionally Additive Models (RAMs)}
\label{alg:ram}
\begin{algorithmic}[1]
\Require Dataset \(\mathcal{D} = \{(\mathbf{x}_i, y_i)\}_{i=1}^n\), black-box model \(f(\mathbf{x})\), regional effect method \(\mathcal{R}\), RAM \(f^{\mathtt{RAM}}(\mathbf{x})\)
\State Fit a black-box model \(f(\mathbf{x})\) to \(\mathcal{D} = \{(\mathbf{x}_i, y_i)\}_{i=1}^n\)
\State Identify the subregions \(\mathcal{R}_{sk}\) using a regional effect method
\State Define the RAM \(f^{\mathtt{RAM}}(\mathbf{x}) = c + \sum_{s=1}^D f^{\mathtt{RAM}}_s(x_s) = c + \sum_{s=1}^D \sum_k f_{sk}(x_s) \when{\xcc \in \mathcal{R}_{sk}}\)
\State Fit the univariate functions \(f_{sk}(x_s)\) for each \(s\) and \(k\).
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{DALE-based Region Detection}
\label{alg:subregion_identification}
\begin{algorithmic}
\Procedure{DALE\_Region\_Detection}{$\mathcal{D}$, $J$, $T$}
    \For {\(s=1, \cdots, D\)}
        \State $\mathtt{D\_before\_split = [D]}$
        \For {\(t=1, \cdots, T\)}
        \EndFor
    \EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{DALE-based Region Detection}
\label{alg:single_feature_subregion}
\begin{algorithmic}
\For{\( c = 1, \cdots, s-1, s+1, \cdots, D\)}
    \State Lalala
    \State $\mathtt{D\_before\_split = [D]}$
    \State $\mathtt{L\_before\_split = [}\mathcal{L}(\mathcal{D}_{s1}) \mathtt{]}$
    \State $\mathtt{L\_after\_split = []}$
\EndFor
\end{algorithmic}
\end{algorithm}

>>>>

% for algorithms
\usepackage[linesnumbered,ruled]{algorithm2e}

\usepackage{natbib}
\bibliographystyle{plainnat}
% set path for bibliography

@@

%\begin{algorithm}
%\caption{Regionally Additive Models (RAMs)}
%\label{alg:ram}
%\begin{algorithmic}[1]
%\Require Dataset \(\mathcal{D} = \{(\mathbf{x}_i, y_i)\}_{i=1}^n\), black-box model \(f(\mathbf{x})\), regional effect method \(\mathcal{R}\), RAM \(f^{\mathtt{RAM}}(\mathbf{x})\)
%\State Fit a black-box model \(f(\mathbf{x})\) to \(\mathcal{D} = \{(\mathbf{x}_i, y_i)\}_{i=1}^n\)
%\State Identify the subregions \(\mathcal{R}_{sk}\) using a regional effect method
%\State Define the RAM \(f^{\mathtt{RAM}}(\mathbf{x}) = c + \sum_{s=1}^D f^{\mathtt{RAM}}_s(x_s) = c + \sum_{s=1}^D \sum_k f_{sk}(x_s) \when{\xcc \in \mathcal{R}_{sk}}\)
%\State Fit the univariate functions \(f_{sk}(x_s)\) for each \(s\) and \(k\).
%\end{algorithmic}
%\end{algorithm}

@@

\begin{algorithm}
\caption{DALE-based Region Detection}
\label{alg:subregion_identification}
\SetAlgoLined
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\BlankLine
\Input{X, J, T}
\BlankLine
\Output{optimal\_splits}
\BlankLine
optimal\_splits[s, c, t] $\gets$ None\;
\For{$s = 1$ to $D$}{
    \For{$c \in \{1, \dots, D\} \setminus \{s\}$}{
        X\_list $\gets [X]$\;
        J\_list $\gets [J]$\;
        \For{$t = 1$ to $T$}{
            best\_split $\gets$ BestSplit(X\_list, J\_list, s)\;
            optimal\_splits[s, c, t] $\gets$ best\_split\;
            X\_list $\gets$ best\_split["X\_list"]\;
            J\_list $\gets$ best\_split["J\_list"]\;
        }
    }
}
\end{algorithm}

\begin{algorithm}
\caption{BestSplit}
\label{alg:single_feature_subregion}
\SetAlgoLined
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}

\Input{X\_list, J\_list, s, c}
\Output{BestSplits}

% Initialize variables
\BlankLine
\uIf{c is categorical}{
    positions $\gets$ GetUniqueValues(X\_list, c)\;
    L $\gets$ []\; % \Comment{List to store interaction values}
    \For{p \textnormal{in} positions}{
        X\_split, J\_split $\gets$ SplitDatasetCategorical(X\_list, J\_list, c, p)\;
        L $\gets$ GetInteraction(X\_split, J\_split)\;
    }
    split\_position = argmin(L)
}
\Else{
    positions $\gets$ GetPosition(X\_list)\;
    L $\gets$ []\; % \Comment{List to store interaction values}
    \For{p \textnormal{in} positions}{
        X\_split, J\_split $\gets$ SplitDatasetNumerical(X\_list, J\_list, c, p)\;
        L $\gets$ GetInteraction(X\_split, J\_split)\;
    }
    split\_position = argmin(L) % \Comment{Store the minimum interaction value for the categorical feature}
}

\Return{BestSplits}
\BlankLine
\end{algorithm}
%
%
%
%\begin{algorithm}
%\caption{BestSplit}
%\label{alg:single_feature_subregion}
%\SetAlgoLined
%\SetKwInOut{Input}{Input}
%\SetKwInOut{Output}{Output}
%
%\Input{X\_list, J\_list, s, t}
%\Output{Splits\_list}
%
%% Initialize variables
%\BlankLine
%\For{c \in \{1, \dots, D\}}{
%    \If{c is categorical}{
%        unique\_values $\gets$ GetUniqueValues(X\_list, c)\;
%        L $\gets$ []\;
%        \For{p in unique\_values}{
%            X\_split, J\_split $\gets$ SplitDatasetCategorical(X\_list, J\_list, c, p)\;
%            L $\gets$ GetInteraction(X\_split, J\_split)\;
%        }
%        L[c] = min(L)\;
%    }
%    \Else{
%        positions $\gets$ GetPosition(X\_list)\;
%        L $\gets$ []\;
%        \For{p in postions}{
%            X\_split, J\_split $\gets$ SplitDatasetNumerical(X\_list, J\_list, c, p)\;
%            L $\gets$ GetInteraction(X\_split, J\_split)\;
%        }
%    }
%}
%\BlankLine
%\end{algorithm}

\begin{algorithm}
\caption{SplitDatasetCategorical}
\label{alg:split_dataset_categorical}
\SetAlgoLined
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\Input{X\_list, J\_list, c, val}
\Output{Split datasets}
X\_split $\gets$ []\;
J\_split $\gets$ []\;
\For{i = 1 to len(X\_list)}{
    ind\_equal $\gets$ X[:, c] $=$ val\;
    ind\_diff $\gets$ X[:, c] $\neq$ val\;
    Append X[ind\_equal], X[ind\_diff] to X\_split\;
    Append J[ind\_equal], J[ind\_diff] to J\_split\;
}
\Return X\_split, J\_split
\end{algorithm}

\begin{algorithm}
\caption{SplitDatasetNumerical}
\label{alg:split_dataset_numerical}
\SetAlgoLined
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\Input{X\_list, J\_list, c, val}
\Output{Split datasets}
X\_split $\gets$ []\;
J\_split $\gets$ []\;
\For{i = 1 to len(X\_list)}{
    ind\_left $\gets$ X[:, c] $\leq$ val\;
    ind\_right $\gets$ X[:, c] $>$ val\;
    Append X[ind\_left], X[ind\_right] to X\_split\;
    Append J[ind\_left], J[ind\_right] to J\_split\;
}
\Return X\_split, J\_split
\end{algorithm}

\begin{algorithm}
\caption{GetInteraction}
\label{alg:get_interaction}
\SetAlgoLined
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\Input{$X\_list$, $J\_list$, $min\_points$}
\Output{weighted average of interaction levels}
$N \gets$ total number of items in $X\_list$\;
$W \gets []$\;
$L \gets []$\;
\For{$i$ in len($X\_list$)}{
    $X \gets X\_list[i]$\;
    $J \gets J\_list[i]$\;
    \uIf{$|X| < min\_points$}{
      Append $\infty$ to $L$\;
    }
    \Else{
        Append $\mathcal{L}(X, J)$ to $L$\;
    }
    Append $\frac{|X|}{N}$ to $W$\;
}
\Return $\text{sum}(L \cdot W)$\;
\end{algorithm}
